The Talking Heads Experiment

### 2015-04-16 chap1/chap1.tex
hyphenation, orphans, widows and some other layout updated

Back in Brussels, an extremely exciting but very 
intense period of research started as I tried to 
apply these principles to language
and subject them to experimental 
scrutiny.\footnote{The earliest papers on these mechanisms are in: \cite{Steels:95b} and 
\cite{Steels:96a}.}
We very quickly built a first 
prototype of a {\it Talking Head} with our own hardware
and low-level software (see \figref{f:plate7}) 
and experimented with language games on mobile 
robotic agents.\footnote{The electronics and tracking software for this 
active camera were built by Tony Belpaeme. The mobile
robot experiments were conducted with Paul Vogt. 
Other early work on the grounding and autonomous acquisition of 
language-like communication systems by robotic systems is 
described in \cite{Steels:97g}. See also: \cite{Billard:1998}.}
New fundamental insights and discoveries emerged almost daily. 
The first experiments, particularly in grounding
language on real robots, were extremely difficult but we 
were clearly making steady progress.

@@

The massive response and thoughtful judgements of the 
public were crucial to validate many aspects of the 
theories put forward in this book. 
But the interactions with a broad public and
the intense discussions that it generated also added new
dimensions to the research. First of all, a whole 
new type of interface between man and machine was
taking shape under our very eyes. In contrast to pre-programmed
computer interfaces, which more often than not make it
difficult to do what one wants, the Talking Heads demonstrated
for the first time the concept of {\itshape negotiated user
interfaces}. The interaction was based on mutual respect
and adaptation of man {\itshape and} machine. Communicative
failure was not fatal but 
an opportunity to fine-tune and negotiate the way communication
would take place in the future.

>>>>

\clearpage
Back in Brussels, an extremely exciting but very 
intense period of research started as I tried to 
apply these principles to language
and subject them to experimental 
scrutiny.\footnote{The earliest papers on these mechanisms are in: \cite{Steels:95b} and 
\cite{Steels:96a}.}
We very quickly built a first 
prototype of a {\it Talking Head} with our own hardware
and low-level software (see \figref{f:plate7}) 
and experimented with language games on mobile 
robotic agents.\footnote{The electronics and tracking software for this 
active camera were built by Tony Belpaeme. The mobile
robot experiments were conducted with Paul Vogt. 
Other early work on the grounding and autonomous acquisition of 
language-like communication systems by robotic systems is 
described in \cite{Steels:97g}. See also: \cite{Billard:1998}.}
New fundamental insights and discoveries emerged almost daily. 
The first experiments, particularly in grounding
language on real robots, were extremely difficult but we 
were clearly making steady progress.

@@

\enlargethispage{1\baselineskip}
The massive response and thoughtful judgements of the 
public were crucial to validate many aspects of the 
theories put forward in this book. 
But the interactions with a broad public and
the intense discussions that it generated also added new
dimensions to the research. First of all, a whole 
new type of interface between man and machine was
taking shape under our very eyes. In contrast to pre-programmed
computer interfaces, which more often than not make it
difficult to do what one wants, the Talking Heads demonstrated
for the first time the concept of {\itshape negotiated user
interfaces}. The interaction was based on mutual respect
and adaptation of man {\itshape and} machine. Communicative
failure was not fatal but 
an opportunity to fine-tune and negotiate the way communication
would take place in the future.

### 2015-04-16 chap10/chap10.tex
hyphenation, orphans, widows and some other layout updated

The Perspective Reversal Experiment worked as follows. A population of \textsc{aibo} robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the `speaker') 
describes the ball-moving event to the other robot (the ``hearer''), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception.pdf}}
\caption{\label{fig:persp-rev-perception}Chain from visual image (left) to an analog world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

@@

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/stages.pdf}}
\caption{\label{fig:stages}Three strategies are compared. In Stage I (left) agents categorise the event from their 
own perspective. In Stage II (middle) they conceptualise the scene from the viewpoint of the other agent as well. In Stage III (right) 
they express that they performed a perspective reversal. Only the third strategy leads to successful performance.} 
\end{figure}

@@

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations.pdf}}
\caption{\label{fig:associations}Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance `(W O) (W U)' is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [AREA v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

Based on these components, I implemented an initial experiment with only two agents.\footnote{These experiments were first reported at the Harvard Evolution of Language conference in 2002 with a paper by Luc Steels
entitled ``Computer Simulations of the origins of Case grammar''. The paper was later deemed 
``too incomprehensible for the Evolution of Language community'' to be acceptable for the conference 
proceedings edited by Maggie Tallerman.}
In 2005 Remi van Trijp picked up this line of research again and worked out the experiment 
further with great skill and sophistication. He extended the 
2-agent simulation to a multi-agent simulation and could finally report a complete running system in his 2008 Ph.D 
thesis \citep{VanTrijp:2014}. Here I give only a brief account of the first attempts towards case grammar
and how this pushed the state of the art in language game research.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/PERACT-setup.pdf}}
\caption{\label{fig:PERACT-setup}The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve.pdf}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill.pdf}}
\caption{\label{fig:adam-eve}Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}

{\bfshape  Scaling up the vision system} 
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called \textsc{peract} \is{\textsc{peract} vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

{\bfshape  Object Identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or ``anchors'' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/PERACT-images.pdf}}
\caption{\label{fig:PERACT-images}Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

{\bfshape  Event Identification} 
The second subsystem detects and tracks events, again at different hierarchical levels. The task is similar
to that of detecting objects but the grouping is based on changes in properties of objects rather than 
on invariances. Event detection is organised in three steps: 
\begin{enumerate}
\item Detect change by computing qualitative descriptions for the movement of an object, the contact between 
two objects, the approaching of two objects, the positioning of an object with respect to another one, etc. 
\item Detect micro-events by grouping image subsequences during which the same configuration 
of qualitative descriptors holds. 
\item Detect events which are defined as sequences of micro-events recognised using probabilistic state machines. 
For example a pick-up event (as in \figref{fig:PERACT-images}) involves three micro-events: (1) the hand 
moves towards the object, (2) the hand touches the object, and (3) both move away together. Processes 
concerned with recovering such events use a library of event definitions which is matched against the stream of micro-events.
\end{enumerate}
An example of output from this subsystem is shown below. Each event has an index (e.g. 77), a time period (e.g. 
from 780 to 803), other events or properties that have played a role in recognising this event, a 
label (e.g. {\verb Appear}), and information of the status (like ongoing or finished).

@@

{\bfshape  Qualitative Descriptors} 
The third subsystem consists of feature detectors that attempt to find qualitative descriptions for units at different levels 
of the object or event hierarchy. The result of all these processes is a set of streams, reporting objects and their 
properties dynamically in response to a changing world as well as the roles these different objects play in the events. 
A (short term) memory of these streams is kept as events unfold. It is used by the semantic system to construct or 
interpret the semantic structures that have to be expressed. Here is a small snapshot of the state of 
this visual memory with time stamps when the states occurred.

\begin{verbatim}
... 
(CAUSE-MOVE-INSIDE ev-7 TRUE) <15:50:2>
(CAUSE-MOVE-INSIDE-TARGET ev-7 obj-1) <15:50:2>
(CAUSE-MOVE-INSIDE-PATIENT ev-7 obj-2) <15:50:2>
(CAUSE-MOVE-INSIDE-AGENT ev-7 YOU) <15:50:2>
(HALT ev-3 TRUE) <15:49:41>
(HALT-ARG ev-3 obj-1) <15:49:41>
(MOVE ev-1 TRUE) <15:49:24>
(MOVE-ARG ev-1 obj-1) <15:49:24>
(STRIPED obj-2) <15:48:42>
(HUGE obj-2) <15:48:42>
(BALL obj-2) <15:48:42>
(WHITE obj-2) <15:48:42>
(OBJECT obj-2) <15:48:42>
(STRIPED obj-1) <15:48:42>
(BIG obj-1) <15:48:42>
(PYRAMID obj-1) <15:48:42>
(YELLOW obj-1) <15:48:42>
(OBJECT obj-1) <15:48:42>
 ...
\end{verbatim}

@@

{\bfshape  Scaling up the language system}\\

Not only the vision system, but also the language system had to be scaled up significantly to do these experiments. 
Together with Nicolas Neubauer, I built upon the formalism I had developed for the early syntax experiments, 
designing and implementing a next generation system, called \textsc{cogar}, \is{\textsc{cogar} (cognitive grammar architecture)} which stands for 
Cognitive Grammar Architecture.\footnote{
No official publication of the \textsc{cogar} system exists. Such technical papers are virtually impossible to get 
published in journals or conferences. There was however a technical report: \cite{Steels:2001}.}
Here is a brief description of this formalism.

@@

Here is an example of a lexical rule that defines the meaning of the word for an event `move-away-from'. The referent is 
the object that is referred to by the word, in this case it is the event itself. 
\begin{verbatim}
((?unit 
   (meaning
      (== (move-away-from ?event ?state)
          (move-away-from-1 ?event ?object)
          (move-away-from-2 ?event ?source)))
   (referent ?event)) 
<-->
((?unit
   (form (== move-away-from)))))
\end{verbatim}
Next, here is an example of a rule for the marker `po'. On the semantic side it ensures that the referent of one unit 
whose referent is ?source is equal to the object filling the argument move-away-from-2 in a move-away-from event. 
On the syntactic side, it triggers when there is a word that has the grammatical category `takes-marker-1' and 
is preceded by the marker `po'. (<< stands for `preceeds') 
\begin{verbatim}
((?predicate 
   (meaning
    (== (move-away-from ?evnt)
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source))))
 (?argument
   (referent ?source))) 
<---> 
((?argument
   (gramcat (== takes-marker-1)))
 (?marker-unit
   (form (== po))) 
(<< ?marker-unit ?argument))))
\end{verbatim}

@@

The marker `po' in the above example is entirely specific for one argument of the move-away-from predicate, as would be the 
case when the second strategy is enacted. To implement the third strategy 
(to achieve more abstract semantic roles) requires that the markers become generic, which implies that 
we get additional rules that recategorise the arguments of a predicate in terms of 
semantic roles like: agent, patient, source, etc. Here is an example of a rule that 
has this effect. role-1 is an example of a semantic role created by the agent as an abstraction of the 
move-away-from-2 argument. 
\begin{verbatim}
((?predicate
   (meaning
    (== (move-away-from ?evnt) 
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source)))))
<---> 
 ((?predicate
    (meaning
      (== (move-away-from ?evnt)
          (move-away-from-1 ?evnt ?patient)
          (move-away-from-2 ?evnt ?source)))) 
    (expanded-meaning (== (role-1 ?evnt ?source))))
\end{verbatim}
And the rules that click eveything together now operate over semantic roles as shown in the following example: 
\begin{verbatim}
 ((?predicate
   (expanded-meaning (== (role-1 ?event ?source))))
  (?unit (referent ?source))) 
<---> 
 ((?unit (gramcat (== takes-marker-1)))
  (?marker-unit (form (== po))) 
 (<< ?marker-unit ?unit))
\end{verbatim}
Recall that all rules must be reversible and that the conditional part must match and the concluding part must unify.\\

{\bfshape  Analogy as the driver of generalisation}\\

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=\textwidth]{chap10/figs/comparison.pdf}}
\caption{\label{fig:comparison}Comparison between two different strategies: The top-graph shows the growth in the average number of 
markers for the event-specific marker strategy. 
The bottom two graphs show the average size of the set of markers when analogy is used. As usual, the y-axis shows 
the number of games played. Overall, fewer markers are needed, making the case marking system more efficient 
and thus explaining why we get abstract semantic roles.}
\end{figure}

@@

>>>>

\enlargethispage{1\baselineskip}
The Perspective Reversal Experiment worked as follows. A population of \textsc{aibo} robots roam around
freely in an unconstrained in-door environment. As soon as one sees the
ball, it comes to a stop and searches for another robot nearby, which also
looks for the ball and stops when it sees it. Then the human
experimenter pushes the ball with a stick so that it rolls a short
distance (see \figref{fig:tokyo-persp}). The two robots next play a description game. One robot (acting as the `speaker') 
describes the ball-moving event to the other robot (the ``hearer''), going through the same semiotic cycle as the Talking 
Heads agents. The game is a success if the meaning conveyed by the speaker is compatible with the hearer's 
own perception of the scene.

@@

\begin{figure}[p]
  \centerline{\includegraphics[width=.90\textwidth]{chap10/figs/persp-rev-perception.pdf}}
\caption{\label{fig:persp-rev-perception}Chain from visual image (left) to an analog world model (middle) and a discrete event description
that is the topic of the language game (right). 
The top row shows this chain for robot A and the bottom row the chain for robot B, both looking at the same ball.}
\end{figure}

@@

\begin{figure}[p]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/stages.pdf}}
\caption{\label{fig:stages}Three strategies are compared. In Stage I (left) agents categorise the event from their 
own perspective. In Stage II (middle) they conceptualise the scene from the viewpoint of the other agent as well. In Stage III (right) 
they express that they performed a perspective reversal. Only the third strategy leads to successful performance.} 
\end{figure}

@@

\begin{figure}[t]
\centerline{
\includegraphics[width=0.75\textwidth]{chap10/figs/associations.pdf}}
\caption{\label{fig:associations}Top: Two lexical associations have become active (association-259 and association-234) to cover the distinctive 
feature-set ([VISIB v-411] [VER v-431]) resulting from discrimination and perception. Bottom: 
The group of words and the meanings match with schemas forming part of a grammatical association that combines the words. }
\end{figure}
An application of these schemas in producing the utterance `(W O) (W U)' is represented in \figref{fig:associations} and shown 
here: 
\begin{verbatim}
127 ++> Speaker: head-40 Topic: 2 Context: 6 5 4 3 1 0
Categorial Perception:
  ([VISIB v-4111] [VER v-431])
  ([VER v-4311] [\textsc{area} v-4061])
Lexicon lookup: (Association-259 Association-234)
Syntactic structure:
  (syn-cat-77
    (syn-slot-50 (syn-cat-75 |(W O)|))
    (syn-slot-51 (syr-cat-76 |(W U)|)))
Semantic structure:
  (sem-cat-78
    (sem-slot-50 (sem-cat-75 (VER v-431)))
    (sem-slot-51 (sem-cat-76 (VISIB v-411))))
Meaning:
  ([VISIB v-4111] [VER v-4311)
Expression: ((W 0) (W U))
\end{verbatim}

@@

Based on these components, I implemented an initial experiment with two agents.\footnote{These experiments were first reported at the Harvard Evolution of Language conference in 2002 with a paper by Luc Steels
entitled ``Computer Simulations of the origins of Case grammar''. The paper was later deemed 
``too incomprehensible for the Evolution of Language community'' to be acceptable for the conference 
proceedings edited by Maggie Tallerman.}
In 2005 Remi van Trijp picked up this line of research again and worked out the experiment 
further with great skill and sophistication. He extended the 
2-agent simulation to a multi-agent simulation and could finally report a complete running system in his 2008 Ph.D 
thesis \citep{VanTrijp:2014}. Here I give only a brief account of the first attempts towards case grammar
and how this pushed the state of the art in language game research.

@@

\begin{figure}[p]
  \centerline{\includegraphics[width=.60\textwidth]{chap10/figs/PERACT-setup.pdf}}
\caption{\label{fig:PERACT-setup}The installation for the case grammar experiments used a more sophisticated vision system and objects that 
could be moved around. We see the two Talking Heads cameras in 
the background. They are oriented towards objects on the table or situations in the 
environment. The monitors show the camera images and the states of image processing, just as for the original 
Talking Heads experiment. 
}
\end{figure}

\begin{figure}[p]
  \centerline{\includegraphics[width=.80\textwidth]{chap10/figs/PERACT-images.pdf}}
\caption{\label{fig:PERACT-images}Top: Some of the images recorded during the action of picking up a cylinder. 
Middle: Results of segmentation algorithms. 
Bottom: The spatio-temporal continuity of objects is tracked. 
}
\end{figure}

@@

\begin{figure}[h]
  \centerline{\includegraphics[width=.40\textwidth]{chap10/figs/adam-eve.pdf}}
\centerline{\includegraphics[width=.70\textwidth]{chap10/figs/jack-give.pdf}}
\centerline{\includegraphics[width=.80\textwidth]{chap10/figs/push-jill.pdf}}
\caption{\label{fig:adam-eve}Top: Computer animation of the speaking and hearing agent used in public demonstrations 
of the case experiment. 
Middle: Example of a puppet scene, where Jill gives a block to Jack. It shows the source image (left) and a (partial) 
result of image processing (right). Bottom: Another example of a puppet scene where Jill pushes the block.}
\end{figure}
\clearpage
\subsubsection{Scaling up the vision system}
The vision system of the Talking Heads only dealt with static scenes. This had to be scaled up to
deal with dynamical scenes and event recognition. It was the focal work of 
Jean-Christophe Baillie at the Sony Computer Science Laboratory in Paris, who built a vision system 
called \textsc{peract} \is{\textsc{peract} vision system} inspired by an event-recognition developed by Jeffrey 
Siskind.\footnote{The vision system is described in the Ph.D thesis of J-C Baillie. A summary is found here: 
\cite{Steels:2003}. Inspiration for the event recognition system came 
primarily from the paper by \cite{Siskind:2000}.}
The vision system was decomposed into three subsystems with information flowing both in a bottom-up
(from image to world model) and top-down way (from world model and predictions to image): the subsystems achieved 
object identification, event identification, and derivation of qualitative descriptors.

\subsubsection{Object Identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or ``anchors'' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\subsubsection{Event Identification} 
The second subsystem detects and tracks events, again at different hierarchical levels. The task is similar
to that of detecting objects but the grouping is based on changes in properties of objects rather than 
on invariances. Event detection is organised in three steps: 
\begin{enumerate}
\item Detect change by computing qualitative descriptions for the movement of an object, the contact between 
two objects, the approaching of two objects, the positioning of an object with respect to another one, etc. 
\item Detect micro-events by grouping image subsequences during which the same configuration 
of qualitative descriptors holds. 
\item Detect events which are defined as sequences of micro-events recognised using probabilistic state machines. 
For example a pick-up event (as in \figref{fig:PERACT-images}) involves three micro-events: (1) the hand 
moves towards the object, (2) the hand touches the object, and (3) both move away together. Processes 
concerned with recovering such events use a library of event definitions which is matched against the stream of micro-events.
\end{enumerate}
An example of output from this subsystem is shown below. Each event has an index (e.g. 77), a time period (e.g. 
from 780 to 803), other events or properties that have played a role in recognising this event, a 
label (e.g. {\verb Appear}), and information of the status (like ongoing or finished).

@@

\subsubsection{Qualitative Descriptors} 
The third subsystem consists of feature detectors that attempt to find qualitative descriptions for units at different levels 
of the object or event hierarchy. The result of all these processes is a set of streams, reporting objects and their 
properties dynamically in response to a changing world as well as the roles these different objects play in the events. 
A (short term) memory of these streams is kept as events unfold. It is used by the semantic system to construct or 
interpret the semantic structures that have to be expressed. Here is a small snapshot of the state of 
this visual memory with time stamps when the states occurred.

\begin{verbatim}
... 
(CAUSE-MOVE-INSIDE ev-7 TRUE) <15:50:2>
(CAUSE-MOVE-INSIDE-TARGET ev-7 obj-1) <15:50:2>
(CAUSE-MOVE-INSIDE-PATIENT ev-7 obj-2) <15:50:2>
(CAUSE-MOVE-INSIDE-AGENT ev-7 YOU) <15:50:2>
(HALT ev-3 TRUE) <15:49:41>
(HALT-ARG ev-3 obj-1) <15:49:41>
(MOVE ev-1 TRUE) <15:49:24>
(MOVE-ARG ev-1 obj-1) <15:49:24>
(STRIPED obj-2) <15:48:42>
(HUGE obj-2) <15:48:42>
(BALL obj-2) <15:48:42>
(WHITE obj-2) <15:48:42>
(OBJECT obj-2) <15:48:42>
(STRIPED obj-1) <15:48:42>
(BIG obj-1) <15:48:42>
(PYRAMID obj-1) <15:48:42>
(\textsc{yellow} obj-1) <15:48:42>
(OBJECT obj-1) <15:48:42>
 ...
\end{verbatim}

@@

\subsubsection{Scaling up the language system}
Not only the vision system, but also the language system had to be scaled up significantly to do these experiments. 
Together with Nicolas Neubauer, I built upon the formalism I had developed for the early syntax experiments, 
designing and implementing a next generation system, called \textsc{cogar}, \is{\textsc{cogar} (cognitive grammar architecture)} which stands for 
Cognitive Grammar Architecture.\footnote{
No official publication of the \textsc{cogar} system exists. Such technical papers are virtually impossible to get 
published in journals or conferences. There was however a technical report: \cite{Steels:2001}.}
Here is a brief description of this formalism.

@@

Here is an example of a lexical rule that defines the meaning of the word for an event `move-away-from'. The referent is 
the object that is referred to by the word, in this case it is the event itself. 
\clearpage
\begin{verbatim}
((?unit 
   (meaning
      (== (move-away-from ?event ?state)
          (move-away-from-1 ?event ?object)
          (move-away-from-2 ?event ?source)))
   (referent ?event)) 
<-->
((?unit
   (form (== move-away-from)))))
\end{verbatim}
Next, here is an example of a rule for the marker `po'. On the semantic side it ensures that the referent of one unit 
whose referent is ?source is equal to the object filling the argument move-away-from-2 in a move-away-from event. 
On the syntactic side, it triggers when there is a word that has the grammatical category `takes-marker-1' and 
is preceded by the marker `po'. (<< stands for `preceeds') 
\begin{verbatim}
((?predicate 
   (meaning
    (== (move-away-from ?evnt)
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source))))
 (?argument
   (referent ?source))) 
<---> 
((?argument
   (gramcat (== takes-marker-1)))
 (?marker-unit
   (form (== po))) 
(<< ?marker-unit ?argument))))
\end{verbatim}

@@

The marker `po' in the above example is entirely specific for one argument of the move-away-from predicate, as would be the 
case when the second strategy is enacted. To implement the third strategy 
(to achieve more abstract semantic roles) requires that the markers become generic, which implies that 
we get additional rules that recategorise the arguments of a predicate in terms of 
semantic roles like: agent, patient, source, etc. Here is an example of a rule that 
has this effect. role-1 is an example of a semantic role created by the agent as an abstraction of the 
move-away-from-2 argument. 
\begin{verbatim}
((?predicate
   (meaning
    (== (move-away-from ?evnt) 
        (move-away-from-1 ?evnt ?patient)
        (move-away-from-2 ?evnt ?source)))))
<---> 
 ((?predicate
    (meaning
      (== (move-away-from ?evnt)
          (move-away-from-1 ?evnt ?patient)
          (move-away-from-2 ?evnt ?source)))) 
    (expanded-meaning (== (role-1 ?evnt ?source))))
\end{verbatim}
And the rules that click eveything together now operate over semantic roles as shown in the following example: 
\begin{verbatim}
 ((?predicate
   (expanded-meaning (== (role-1 ?event ?source))))
  (?unit (referent ?source))) 
<---> 
 ((?unit (gramcat (== takes-marker-1)))
  (?marker-unit (form (== po))) 
 (<< ?marker-unit ?unit))
\end{verbatim}
Recall that all rules must be reversible and that the conditional part must match and the concluding part must unify.

\subsubsection{Analogy as the driver of generalisation}

@@

@@

\begin{figure}[h]
  \centerline{\includegraphics[width=\textwidth]{chap10/figs/comparison.pdf}}
\caption{\label{fig:comparison}Comparison between two different strategies: The top-graph shows the growth in the average number of 
markers for the event-specific marker strategy. 
The bottom two graphs show the average size of the set of markers when analogy is used. As usual, the y-axis shows 
the number of games played. Overall, fewer markers are needed, making the case marking system more efficient 
and thus explaining why we get abstract semantic roles.}
\end{figure}
